\documentclass[]{article}
\usepackage[linesnumbered,boxed]{algorithm2e}
\usepackage{amsmath}
\usepackage{graphicx}
%opening
\title{}
\author{}

\begin{document}
\title{Chapter 2 solution}
\maketitle
\begin{enumerate}
\item[3.1-1] Since f(n) and g(n) are both nonnegative\\
$ \exists c1=1>0, k1>0, \forall n>k1, \max(f(n),g(n)<=f(n)+g(n) $\\
$ \exists c2=0.5>0, k2>0, \forall n>k2, \max(f(n),g(n)>=0.5*(f(n)+g(n)) $\\
Therefore, $ \max(f(n),g(n)) = \Theta(f(n)+g(n)) $

\item[3.1-2] $ \frac{(n+a)^b}{n^b} = (1+\frac{a}{n})^b \in (1/2, 2)\ \  \text{as}\ \  n \to \infty$

Thus $(n+a)^b = \Theta(n^b)$

\item[3.1-3] It is meaningless because the $O$ notation is a upper bound, while "At least" is supposed to demonstrate some lower bound of the running time.

\item[3.1-4] $2^{n+1} = O(2^n)$ since $2^{n+1} < 3*2^{n}$\\
$2^{2n} \neq O(2^n)$ since $\frac{2^{2n}}{2^n} = 2^n \ \ \to \ \ \infty$ as n goes to infinity

\item[3.1-5] Prove that $f(n) = \Theta(g(n))$ if and only if\\ $f(n) = O(g(n))$ and $ f(n) = \Omega(g(n)) $

Just rewrite the definitions and it is easy to find how this relationship holds.

\item[3.1-6] If and algorithm has running time $\Theta(g(n))$, then it must be both $O(g(n))$ and $\Omega(g(n))$\\
Since it includes the worst case and the best case, then it is trivial that $ T_{worst} = O(g(n)) $ and $T_{best} = \Omega(g(n))$

\item[3.1-7] Prove that $o(g(n) \cap \omega(g(n)) = \emptyset $\\
Suppose there is a function $f(n)$ in this intersection set. Then by the property of these two notations, we can see that $$ \lim\limits_{n\to \infty} \frac{f(n)}{g(n)} = 0\ \  \wedge  \ \ \lim\limits_{n\to \infty}\frac{f(n)}{g(n)} > 0 $$
which are mutually exclusive. 

\item[3.1-8] Definition of $\Omega(g(n,m))$\\
$ \Omega(g(n,m)) = \\ \{ f(n,m): \exists \text{c,n0,m0$>$0 s.t.} f(n,m)>=cg(n,m) \forall n>=n0 \wedge m>=m0 \}$

$\Theta(g(n,m)) = \Omega(g(n,m)) \cap O(g(n,m)) $

\item[3.2-1] Proof: \\
By given condition, we know that\\ $\forall x>y,\ \ f(x)>f(y)$ and $\forall x>y,\ \ f(x)>f(y)$\\
similar for g(n)

Then it is easy to see that\\ $ \forall x>y, f(x)+g(x)>f(y)+g(y) $\\
$ \forall x>y,\ \ g(x)>g(y)\implies f(g(x))>f(g(y)) $\\
If f and g are both nonnegative, then their product will be trivially increasing.

\item[3.2-2] Prove that $ a^{\log_b(c)} = c^{\log_b(a)} $\\
Suppose the equation holds. Take log a on both sides of the equation, we get $$ \log_b(c) = \log_b(a) \log_a(c) $$

Decompose the log terms with the relation that $ \log_x(y) = \frac{\lg x}{\lg y} $, then it is easy to see that the left hand side and the right side are equivalent.

Since the process above is revertible, it can be deduced that the original equation holds.

\item[3.2-3] Prove that $\lg(n!) = \Theta(n\lg n)$\\
Proof: $\frac{\lg(n!)}{n\lg(n)} = \frac{\lg(n!)}{\lg(n^n)}$\\

It is easy to say that this term is greater than 0 and is upper-bounded by 1. So the O relation is obvious, now we need to show that this term doesn't converge to 0. 

By Stirling Approximation, 
$$
n! = \sqrt{2\pi n}(\frac{ n}{e})^{n}\Theta(1)
$$

$$ \lg(n!) = 1/2*\lg n + \Theta(\lg(n^n)) = \Theta(\lg(n^n))$$


The sum excluding $f(k)$ in the first equation gives $ \Theta(\lg m) $ which is equivalent to $ \Theta(n) $ 



Then it is trivial to see that $ \lg(n!) = \Theta(n\lg n) $

Prove: $ n! = \omega(2^n) $\\
Proof: it is true that $ 4! > 2^4 $
Starting from $n=4$, suppose that $ \frac{n!}{2^n} $ converges to some constant c (say equals $ c-e $ for some infinitesimally small $ e $) when n is greater than k. We can always find some n such that$(n+1)/2 > \frac{c+1}{c-e}$ such that $\frac{(n+1)!}{2^{n+1}} > c$ which contradicts the condition for converging. We can deduce that there can't be upper bound for the result of this division and then we can state that 
$$
n! = \omega(2^n)
$$

Prove: $ n! = o(n^n) $\\
This can be proved by deducing that $\frac{n!}{n^n}$ is monotonically decreasing and it is bounded by 0 and 1.

\item[3.2-4] Is the function $ \lceil \lg(n) \rceil ! $ polynomially bounded?\\
No this is not polynomially bounded. Recall the result we have proved in 3.2-3, $$ \lg( \lceil \lg(n) \rceil ! ) = \omega( 2^{\lceil \lg(n) \rceil} ) = \omega (n) $$

However, for any polynomial P(n), $$ \lg(P(n)) = \Theta(n) $$

Thus it is obvious that this function is asymptotically larger than any polynomial.

Is the function $ \lceil \lg \lg(n) \rceil ! $ polynomially bounded?\\

Yes, this function is polynomially bounded. By Stirling's approximation, 
$$
\lceil\lg\lg n\rceil! = t! = \sqrt{2\pi t}(\frac{t}{e})^{t}(1+\Theta(1/t))
$$

$$ \lg (t!) = O(t \lg (t) ) = O(\lceil\lg\lg n\rceil \lceil\lg\lg\lg n\rceil) = O(2^{\lceil\lg\lg n\rceil}) = O(\lg n)$$

while $$ \lg P(n) = \Theta(n)  $$

Thus it is polynomially bounded.
\end{enumerate}
\end{document}